---
title: "meta"
output: html_document
---

# install packages

```{r}
install.packages("readxl")
install.packages("metafor")
library(readxl)
library(dplyr)
library(ggplot2)
library(broom)
library(readr)
library(metafor)
```

# load dataset

```{r}
NBA <- read_excel("/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/dat/NBA_noTUEPAC_IPD.xlsx")
```

```{r}
NBA_EUR <- NBA %>%
  filter(ANCESTRY == "EUR")
```

```{r}
# process the dataset
# convert NA into real missing values
NBA_EUR <- NBA_EUR %>%
  mutate(across(
    everything(),
    ~ {
      x <- as.character(.x)
      x <- trimws(x)
      x[x %in% c("", "NA", "Na", "N/A", "NULL", "null")] <- NA
      x
    }
  ))

NBA_EUR$study <- as.factor(NBA_EUR$study)
```

```{r}
NBA_EUR$PRS <- as.numeric(NBA_EUR$PRS)
NBA_EUR$PC1 <- as.numeric(NBA_EUR$PC1)
NBA_EUR$PC2 <- as.numeric(NBA_EUR$PC2)
NBA_EUR$PC3 <- as.numeric(NBA_EUR$PC3)
NBA_EUR$PC4 <- as.numeric(NBA_EUR$PC4)
NBA_EUR$PC5 <- as.numeric(NBA_EUR$PC5)
NBA_EUR$age_of_onset <- as.numeric(NBA_EUR$age_of_onset)
NBA_EUR$mds_updrs_part_i_summary_score <- as.numeric(NBA_EUR$mds_updrs_part_i_summary_score)
NBA_EUR$AAV_UPDRSI <- as.numeric(NBA_EUR$AAV_UPDRSI)
NBA_EUR$mds_updrs_part_ii_summary_score <- as.numeric(NBA_EUR$mds_updrs_part_ii_summary_score)
NBA_EUR$AAV_UPDRSII <- as.numeric(NBA_EUR$AAV_UPDRSII)
```

# sample size in each cohort for each variable

```{r}
df <- NBA_EUR %>%
  filter(!is.na(age_of_onset))

df <- NBA_EUR %>%
  filter(!is.na(mds_updrs_part_i_summary_score))

df <- NBA_EUR %>%
  filter(!is.na(mds_updrs_part_ii_summary_score))

df <- NBA_EUR %>%
  filter(!is.na(mds_updrs_part_iii_summary_score))

df <- NBA_EUR %>%
  filter(!is.na(mds_updrs_part_iv_summary_score))
```

```{r}
df <- NBA_EUR %>%
  filter(!is.na(COGNITION))

df <- NBA_EUR %>%
  filter(!is.na(depress_test_score))

df <- NBA_EUR %>%
  filter(!is.na(gds15_total_score))

df <- NBA_EUR %>%
  filter(!is.na(rbd_summary_score))

df <- NBA_EUR %>%
  filter(!is.na(ess_total_score))
```

# SEX regression in each cohort

```{r}
# create a new column male = 1, female = 2
NBA_EUR$SEX = factor(NBA_EUR$SEX, levels = c(1, 2), labels = c("male", "female"))
# linear regression in 28 cohorts separately

study_keep <- c(
  "APGS", "COPN", "CORIELL", "DENOPA", "DODONA","IDIBAPS","LARGEPD","LUXPARK", "MDGAP-IBB","MDGAP-QSBB", "NZPEGS","PAGE","PDGEN","PDGNRTN","PPMI-G","PPMI-N","PROMO-EPIPARK","PROMO-GENEPARK","PROMO-ROPAD","SP-SEV","STELLENBOS","UPENNU19","VIPD")

NBA_EUR_sex <- NBA_EUR %>%
  filter(
    !is.na(SEX),
    SEX != "",
    PHENO != "1"
  )

lm_SEX <- NBA_EUR_sex %>%
  filter(study %in% study_keep) %>%
  group_by(study) %>%
  do(tidy(lm(PRS ~ SEX + PC1 + PC2 + PC3 + PC4 + PC5, data = .))) %>%
  ungroup()
```

## meta analysis

```{r}
sex_meta <- lm_SEX %>%
  filter(term == "SEXfemale") %>%
  select(study, estimate, std.error, p.value)

# add 95%CI column
sex_meta <- sex_meta %>%
  mutate(
    ci.lb = estimate - 1.96 * std.error,
    ci.ub = estimate + 1.96 * std.error
  )

# Fixed-effect model
fixed <- rma.uni(yi = estimate,
                     sei = std.error,
                     data = sex_meta,
                     method = "FE")

# Random-effects model
random <- rma.uni(yi = estimate,
                      sei = std.error,
                      data = sex_meta,
                      method = "REML")

writeLines(capture.output(summary(fixed)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/sex_fixed.txt")
writeLines(capture.output(summary(random)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/sex_random.txt")
```

```{r}
random_row <- tibble(
  study = "Random-effects Model",
  estimate = as.numeric(random$b),
  std.error = random$se,
  p.value = random$pval,
  ci.lb = random$ci.lb,
  ci.ub = random$ci.ub
)
sex_meta <- bind_rows(sex_meta, random_row)

fixed_row <- tibble(
  study = "Fixed-effects Model",
  estimate = as.numeric(fixed$b),
  std.error = fixed$se,
  p.value = fixed$pval,
  ci.lb = fixed$ci.lb,
  ci.ub = fixed$ci.ub
)
sex_meta <- bind_rows(sex_meta, fixed_row)

sex_meta_t <- sex_meta %>%
  select(study, estimate, ci.lb, ci.ub, p.value) %>%
  column_to_rownames("study") %>%
  t() %>%
  as.data.frame()

write.csv(sex_meta, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/sex_meta.csv")
write.csv(sex_meta_t, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/sex_meta_t.csv")
```

## visualization with forest plot

# age at onset regression in each cohort

```{r}
# linear regression in 28 cohorts separately

study_keep <- c(
  "APGS", "CORIELL", "DODONA", "PAGE", "PDGNRTN",
  "PROMO-EPIPARK", "SP-SEV", "VIPD", "COPN", "DENOPA", "IDIBAPS", "MDGAP-QSBB", "NZPEGS", 
  "PDGEN", "PPMI-N", "PROMO-GENEPARK", "STELLENBOS","UPENNU19"
)

lm_AAO <- NBA_EUR %>%
  filter(study %in% study_keep) %>%
  group_by(study) %>%
  do(tidy(lm(age_of_onset ~ PRS + SEX + PC1 + PC2 + PC3 + PC4 + PC5, data = .))) %>%
  ungroup()
```

## meta analysis

```{r}
AAO_meta <- lm_AAO %>%
  filter(term == "PRS") %>%
  select(study, estimate, std.error, p.value)

# add 95%CI column
AAO_meta <- AAO_meta %>%
  mutate(
    ci.lb = estimate - 1.96 * std.error,
    ci.ub = estimate + 1.96 * std.error
  )

# Fixed-effect model
fixed <- rma.uni(yi = estimate,
                     sei = std.error,
                     data = AAO_meta,
                     method = "FE")
# Random-effects model
random <- rma.uni(yi = estimate,
                      sei = std.error,
                      data = AAO_meta,
                      method = "REML")

writeLines(capture.output(summary(fixed)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/aao_fixed.txt")
writeLines(capture.output(summary(random)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/aao_random.txt")
```

```{r}
random_row <- tibble(
  study = "Random-effects Model",
  estimate = as.numeric(random$b),
  std.error = random$se,
  p.value = random$pval,
  ci.lb = random$ci.lb,
  ci.ub = random$ci.ub
)
AAO_meta <- bind_rows(AAO_meta, random_row)

fixed_row <- tibble(
  study = "Fixed-effects Model",
  estimate = as.numeric(fixed$b),
  std.error = fixed$se,
  p.value = fixed$pval,
  ci.lb = fixed$ci.lb,
  ci.ub = fixed$ci.ub
)
AAO_meta <- bind_rows(AAO_meta, fixed_row)

AAO_meta_t <- AAO_meta %>%
  select(study, estimate, ci.lb, ci.ub, p.value) %>%
  column_to_rownames("study") %>%
  t() %>%
  as.data.frame()

write.csv(AAO_meta, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/AAO_meta.csv")
write.csv(AAO_meta_t, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/AAO_meta_t.csv")
```

# UPDRS-I regression in each cohort

```{r}
study_keep <- c(
  "PPMI-N", "COPN", "PPMI-G"
)

lm_UPDRSI <- NBA_EUR %>%
  filter(study %in% study_keep) %>%
  group_by(study) %>%
  do(tidy(lm(mds_updrs_part_i_summary_score ~ PRS + AAV_UPDRSI + SEX + PC1 + PC2 + PC3 + PC4 + PC5, data = .))) %>%
  ungroup()
```

## meta analysis

```{r}
updrsi_meta <- lm_UPDRSI %>%
  filter(term == "PRS") %>%
  select(study, estimate, std.error, p.value)

# add 95%CI column
updrsi_meta <- updrsi_meta %>%
  mutate(
    ci.lb = estimate - 1.96 * std.error,
    ci.ub = estimate + 1.96 * std.error
  )

# Fixed-effect model
fixed <- rma.uni(yi = estimate,
                     sei = std.error,
                     data = updrsi_meta,
                     method = "FE")

# Random-effects model
random <- rma.uni(yi = estimate,
                      sei = std.error,
                      data = updrsi_meta,
                      method = "REML")

writeLines(capture.output(summary(fixed)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/updrsi_fixed.txt")
writeLines(capture.output(summary(random)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/updrsi_random.txt")
```

```{r}
random_row <- tibble(
  study = "Random-effects Model",
  estimate = as.numeric(random$b),
  std.error = random$se,
  p.value = random$pval,
  ci.lb = random$ci.lb,
  ci.ub = random$ci.ub
)
updrsi_meta <- bind_rows(updrsi_meta, random_row)

fixed_row <- tibble(
  study = "Fixed-effects Model",
  estimate = as.numeric(fixed$b),
  std.error = fixed$se,
  p.value = fixed$pval,
  ci.lb = fixed$ci.lb,
  ci.ub = fixed$ci.ub
)
updrsi_meta <- bind_rows(updrsi_meta, fixed_row)

updrsi_meta_t <- updrsi_meta %>%
  select(study, estimate, ci.lb, ci.ub, p.value) %>%
  column_to_rownames("study") %>%
  t() %>%
  as.data.frame()

write.csv(updrsi_meta, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/updrsi_meta.csv")
write.csv(updrsi_meta_t, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/updrsi_meta_t.csv")
```

# UPDRSII regression in each cohort

```{r}
study_keep <- c(
  "PPMI-N", "COPN", "PPMI-G")

lm_updrsii <- NBA_EUR %>%
  filter(study %in% study_keep) %>%
  group_by(study) %>%
  do(tidy(lm(mds_updrs_part_ii_summary_score ~ PRS + AAV_UPDRSII + SEX + PC1 + PC2 + PC3 + PC4 + PC5, data = .))) %>%
  ungroup()
```

## meta analysis

```{r}
updrsii_meta <- lm_updrsii %>%
  filter(term == "PRS") %>%
  select(study, estimate, std.error, p.value)

# add 95%CI column
updrsii_meta <- updrsii_meta %>%
  mutate(
    ci.lb = estimate - 1.96 * std.error,
    ci.ub = estimate + 1.96 * std.error
  )

# Fixed-effect model
fixed <- rma.uni(yi = estimate,
                     sei = std.error,
                     data = updrsii_meta,
                     method = "FE")

# Random-effects model
random <- rma.uni(yi = estimate,
                      sei = std.error,
                      data = updrsii_meta,
                      method = "REML")

writeLines(capture.output(summary(fixed)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/updrsii_fixed.txt")
writeLines(capture.output(summary(random)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/updrsii_random.txt")
```

```{r}
random_row <- tibble(
  study = "Random-effects Model",
  estimate = as.numeric(random$b),
  std.error = random$se,
  p.value = random$pval,
  ci.lb = random$ci.lb,
  ci.ub = random$ci.ub
)
updrsii_meta <- bind_rows(updrsii_meta, random_row)

fixed_row <- tibble(
  study = "Fixed-effects Model",
  estimate = as.numeric(fixed$b),
  std.error = fixed$se,
  p.value = fixed$pval,
  ci.lb = fixed$ci.lb,
  ci.ub = fixed$ci.ub
)
updrsii_meta <- bind_rows(updrsii_meta, fixed_row)

updrsii_meta_t <- updrsii_meta %>%
  select(study, estimate, ci.lb, ci.ub, p.value) %>%
  column_to_rownames("study") %>%
  t() %>%
  as.data.frame()

write.csv(updrsii_meta, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/updrsii_meta.csv")
write.csv(updrsii_meta_t, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/updrsii_meta_t.csv")
```

# updrs-iii regression in each cohort

```{r}
NBA_EUR$mds_updrs_part_iii_summary_score <- as.numeric(NBA_EUR$mds_updrs_part_iii_summary_score)
NBA_EUR$AAV_UPDRSIII <- as.numeric(NBA_EUR$AAV_UPDRSIII)
```

```{r}
# linear regression in 4 cohorts separately

study_keep <- c(
  "BIOFIND", "PDGNRTN", "PPMI-N", "COPN", "PPMI-G"
)

lm_updrsiii <- NBA_EUR %>%
  filter(study %in% study_keep) %>%
  group_by(study) %>%
  do(tidy(lm(mds_updrs_part_iii_summary_score ~ PRS + AAV_UPDRSIII + SEX + PC1 + PC2 + PC3 + PC4 + PC5, data = .))) %>%
  ungroup()
```

## meta analysis

```{r}
updrsiii_meta <- lm_updrsiii %>%
  filter(term == "PRS") %>%
  select(study, estimate, std.error, p.value)

# add 95%CI column
updrsiii_meta <- updrsiii_meta %>%
  mutate(
    ci.lb = estimate - 1.96 * std.error,
    ci.ub = estimate + 1.96 * std.error
  )

# Fixed-effect model
fixed <- rma.uni(yi = estimate,
                     sei = std.error,
                     data = updrsiii_meta,
                     method = "FE")

# Random-effects model
random <- rma.uni(yi = estimate,
                      sei = std.error,
                      data = updrsiii_meta,
                      method = "REML")
					  
writeLines(capture.output(summary(fixed)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/updrsiii_fixed.txt")
writeLines(capture.output(summary(random)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/updrsiii_random.txt")
```

```{r}
random_row <- tibble(
  study = "Random-effects Model",
  estimate = as.numeric(random$b),
  std.error = random$se,
  p.value = random$pval,
  ci.lb = random$ci.lb,
  ci.ub = random$ci.ub
)
updrsiii_meta <- bind_rows(updrsiii_meta, random_row)

fixed_row <- tibble(
  study = "Fixed-effects Model",
  estimate = as.numeric(fixed$b),
  std.error = fixed$se,
  p.value = fixed$pval,
  ci.lb = fixed$ci.lb,
  ci.ub = fixed$ci.ub
)
updrsiii_meta <- bind_rows(updrsiii_meta, fixed_row)

updrsiii_meta_t <- updrsiii_meta %>%
  select(study, estimate, ci.lb, ci.ub, p.value) %>%
  column_to_rownames("study") %>%
  t() %>%
  as.data.frame()

write.csv(updrsiii_meta, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/updrsiii_meta.csv")
write.csv(updrsiii_meta_t, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/updrsiii_meta_t.csv")
```

# UPDRS-IV regression in each cohort

```{r}
NBA_EUR$mds_updrs_part_iv_summary_score <- as.numeric(NBA_EUR$mds_updrs_part_iv_summary_score)
NBA_EUR$AAV_UPDRSIV <- as.numeric(NBA_EUR$AAV_UPDRSIV)
```

```{r}
# linear regression in 2 cohorts separately
study_keep <- c(
  "PPMI-N", "COPN","PPMI-G"
)

lm_updrsiv <- NBA_EUR %>%
  filter(study %in% study_keep) %>%
  group_by(study) %>%
  do(tidy(lm(mds_updrs_part_iv_summary_score ~ PRS + AAV_UPDRSIV + SEX + PC1 + PC2 + PC3 + PC4 + PC5, data = .))) %>%
  ungroup()
```

## meta analysis

```{r}
updrsiv_meta <- lm_updrsiv %>%
  filter(term == "PRS") %>%
  select(study, estimate, std.error, p.value)

# add 95%CI column
updrsiv_meta <- updrsiv_meta %>%
  mutate(
    ci.lb = estimate - 1.96 * std.error,
    ci.ub = estimate + 1.96 * std.error
  )

# Fixed-effect model
fixed <- rma.uni(yi = estimate,
                     sei = std.error,
                     data = updrsiv_meta,
                     method = "FE")

# Random-effects model
random <- rma.uni(yi = estimate,
                      sei = std.error,
                      data = updrsiv_meta,
                      method = "REML")
					  
writeLines(capture.output(summary(fixed)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/updrsiv_fixed.txt")
writeLines(capture.output(summary(random)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/updrsiv_random.txt")
```

```{r}
random_row <- tibble(
  study = "Random-effects Model",
  estimate = as.numeric(random$b),
  std.error = random$se,
  p.value = random$pval,
  ci.lb = random$ci.lb,
  ci.ub = random$ci.ub
)
updrsiv_meta <- bind_rows(updrsiv_meta, random_row)

fixed_row <- tibble(
  study = "Fixed-effects Model",
  estimate = as.numeric(fixed$b),
  std.error = fixed$se,
  p.value = fixed$pval,
  ci.lb = fixed$ci.lb,
  ci.ub = fixed$ci.ub
)
updrsiv_meta <- bind_rows(updrsiv_meta, fixed_row)

updrsiv_meta_t <- updrsiv_meta %>%
  select(study, estimate, ci.lb, ci.ub, p.value) %>%
  column_to_rownames("study") %>%
  t() %>%
  as.data.frame()

write.csv(updrsiv_meta, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/updrsiv_meta.csv")
write.csv(updrsiv_meta_t, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/updrsiv_meta_t.csv")
```

# cognition regression in each cohort

```{r}
NBA_EUR$COGNITION <- as.numeric(NBA_EUR$COGNITION)
NBA_EUR$AAV_COGNITION <- as.numeric(NBA_EUR$AAV_COGNITION)
```

```{r}

NBA_EUR <- NBA_EUR %>%
  mutate(
    COGNITION = as.numeric(COGNITION)
  )
# linear regression in 4 cohorts separately

study_keep <- c(
  "CORIELL", "PROMO-EPIPARK", "SP-SEV", "COPN", "BIOFIND", "DODONA"
)

lm_cog <- NBA_EUR %>%
  filter(study %in% study_keep) %>%
  group_by(study) %>%
  do(tidy(glm(COGNITION ~ PRS + AAV_COGNITION + SEX + PC1 + PC2 + PC3 + PC4 + PC5, data = ., family = binomial, na.action = na.omit))) %>%
  ungroup()
```

## meta analysis

```{r}
cog_meta <- lm_cog %>%
  filter(term == "PRS") %>%
  select(study, estimate, std.error, p.value)

# add 95%CI column
cog_meta <- cog_meta %>%
  mutate(
    ci.lb = estimate - 1.96 * std.error,
    ci.ub = estimate + 1.96 * std.error
  )

# Fixed-effect model
fixed <- rma.uni(yi = estimate,
                     sei = std.error,
                     data = cog_meta,
                     method = "FE")

# Random-effects model
random <- rma.uni(yi = estimate,
                      sei = std.error,
                      data = cog_meta,
                      method = "REML")
					  
writeLines(capture.output(summary(fixed)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/cog_fixed.txt")
writeLines(capture.output(summary(random)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/cog_random.txt")
```

```{r}
random_row <- tibble(
  study = "Random-effects Model",
  estimate = as.numeric(random$b),
  std.error = random$se,
  p.value = random$pval,
  ci.lb = random$ci.lb,
  ci.ub = random$ci.ub
)
cog_meta <- bind_rows(cog_meta, random_row)

fixed_row <- tibble(
  study = "Fixed-effects Model",
  estimate = as.numeric(fixed$b),
  std.error = fixed$se,
  p.value = fixed$pval,
  ci.lb = fixed$ci.lb,
  ci.ub = fixed$ci.ub
)
cog_meta <- bind_rows(cog_meta, fixed_row)

cog_meta_t <- cog_meta %>%
  select(study, estimate, ci.lb, ci.ub, p.value) %>%
  column_to_rownames("study") %>%
  t() %>%
  as.data.frame()

write.csv(cog_meta, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/cog_meta.csv")
write.csv(cog_meta_t, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/cog_meta_t.csv")
```

# depression regression in each cohort

```{r}
NBA_EUR$depress_test_score <- as.numeric(NBA_EUR$depress_test_score)
NBA_EUR$AAV_DEPRESS <- as.numeric(NBA_EUR$AAV_DEPRESS)
```

```{r}
# linear regression in 3 cohorts separately

study_keep <- c(
  "PROMO-EPIPARK", "PROMO-SFB134", "CORIELL"
)

lm_depress <- NBA_EUR %>%
  filter(study %in% study_keep) %>%
  group_by(study) %>%
  do(tidy(lm(depress_test_score ~ PRS + AAV_DEPRESS + SEX + PC1 + PC2 + PC3 + PC4 + PC5, data = .))) %>%
  ungroup()
```

## meta analysis

```{r}
depress_meta <- lm_depress %>%
  filter(term == "PRS") %>%
  select(study, estimate, std.error, p.value)

# add 95%CI column
depress_meta <- depress_meta %>%
  mutate(
    ci.lb = estimate - 1.96 * std.error,
    ci.ub = estimate + 1.96 * std.error
  )

# Fixed-effect model
fixed <- rma.uni(yi = estimate,
                     sei = std.error,
                     data = depress_meta,
                     method = "FE")

# Random-effects model
random <- rma.uni(yi = estimate,
                      sei = std.error,
                      data = depress_meta,
                      method = "REML")
					  
writeLines(capture.output(summary(fixed)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/depress_fixed.txt")
writeLines(capture.output(summary(random)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/depress_random.txt")
```

```{r}
random_row <- tibble(
  study = "Random-effects Model",
  estimate = as.numeric(random$b),
  std.error = random$se,
  p.value = random$pval,
  ci.lb = random$ci.lb,
  ci.ub = random$ci.ub
)
depress_meta <- bind_rows(depress_meta, random_row)

fixed_row <- tibble(
  study = "Fixed-effects Model",
  estimate = as.numeric(fixed$b),
  std.error = fixed$se,
  p.value = fixed$pval,
  ci.lb = fixed$ci.lb,
  ci.ub = fixed$ci.ub
)
depress_meta <- bind_rows(depress_meta, fixed_row)

depress_meta_t <- depress_meta %>%
  select(study, estimate, ci.lb, ci.ub, p.value) %>%
  column_to_rownames("study") %>%
  t() %>%
  as.data.frame()

write.csv(depress_meta, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/depress_meta.csv")
write.csv(depress_meta_t, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/depress_meta_t.csv")

```

# GDS regression in each cohort

```{r}
NBA_EUR$gds15_total_score <- as.numeric(NBA_EUR$gds15_total_score)
NBA_EUR$AAV_GDS <- as.numeric(NBA_EUR$AAV_GDS)
```

```{r}
# linear regression in 2 cohorts separately

study_keep <- c(
  "PPMI-N", "PPMI-G"
)

lm_gds <- NBA_EUR %>%
  filter(study %in% study_keep) %>%
  group_by(study) %>%
  do(tidy(lm(gds15_total_score ~ PRS + AAV_GDS + SEX + PC1 + PC2 + PC3 + PC4 + PC5, data = .))) %>%
  ungroup()
```

## meta analysis

```{r}
gds_meta <- lm_gds %>%
  filter(term == "PRS") %>%
  select(study, estimate, std.error, p.value)

# add 95%CI column
gds_meta <- gds_meta %>%
  mutate(
    ci.lb = estimate - 1.96 * std.error,
    ci.ub = estimate + 1.96 * std.error
  )

# Fixed-effect model
fixed <- rma.uni(yi = estimate,
                     sei = std.error,
                     data = gds_meta,
                     method = "FE")

# Random-effects model
random <- rma.uni(yi = estimate,
                      sei = std.error,
                      data = gds_meta,
                      method = "REML")

writeLines(capture.output(summary(fixed)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/gds_fixed.txt")
writeLines(capture.output(summary(random)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/gds_random.txt")
```

```{r}
random_row <- tibble(
  study = "Random-effects Model",
  estimate = as.numeric(random$b),
  std.error = random$se,
  p.value = random$pval,
  ci.lb = random$ci.lb,
  ci.ub = random$ci.ub
)
gds_meta <- bind_rows(gds_meta, random_row)

fixed_row <- tibble(
  study = "Fixed-effects Model",
  estimate = as.numeric(fixed$b),
  std.error = fixed$se,
  p.value = fixed$pval,
  ci.lb = fixed$ci.lb,
  ci.ub = fixed$ci.ub
)
gds_meta <- bind_rows(gds_meta, fixed_row)

gds_meta_t <- gds_meta %>%
  select(study, estimate, ci.lb, ci.ub, p.value) %>%
  column_to_rownames("study") %>%
  t() %>%
  as.data.frame()

write.csv(gds_meta, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/gds_meta.csv")
write.csv(gds_meta_t, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/gds_meta_t.csv")

```

# RBDSQ regression in each cohort

```{r}
NBA_EUR$rbd_summary_score <- as.numeric(NBA_EUR$rbd_summary_score)
NBA_EUR$AAV_RBD <- as.numeric(NBA_EUR$AAV_RBD)
```

```{r}
# linear regression in 2 cohorts separately

study_keep <- c(
  "PPMI-N", "PPMI-G"
)

lm_rbd <- NBA_EUR %>%
  filter(study %in% study_keep) %>%
  group_by(study) %>%
  do(tidy(lm(rbd_summary_score ~ PRS + AAV_RBD + SEX + PC1 + PC2 + PC3 + PC4 + PC5, data = .))) %>%
  ungroup()
```

# meta analysis

```{r}
rbd_meta <- lm_rbd %>%
  filter(term == "PRS") %>%
  select(study, estimate, std.error, p.value)

# add 95%CI column
rbd_meta <- rbd_meta %>%
  mutate(
    ci.lb = estimate - 1.96 * std.error,
    ci.ub = estimate + 1.96 * std.error
  )

# Fixed-effect model
fixed <- rma.uni(yi = estimate,
                     sei = std.error,
                     data = rbd_meta,
                     method = "FE")

# Random-effects model
random <- rma.uni(yi = estimate,
                      sei = std.error,
                      data = rbd_meta,
                      method = "REML")
					  
writeLines(capture.output(summary(fixed)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/rbd_fixed.txt")
writeLines(capture.output(summary(random)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/rbd_random.txt")

```

```{r}
random_row <- tibble(
  study = "Random-effects Model",
  estimate = as.numeric(random$b),
  std.error = random$se,
  p.value = random$pval,
  ci.lb = random$ci.lb,
  ci.ub = random$ci.ub
)
rbd_meta <- bind_rows(rbd_meta, random_row)

fixed_row <- tibble(
  study = "Fixed-effects Model",
  estimate = as.numeric(fixed$b),
  std.error = fixed$se,
  p.value = fixed$pval,
  ci.lb = fixed$ci.lb,
  ci.ub = fixed$ci.ub
)
rbd_meta <- bind_rows(rbd_meta, fixed_row)

rbd_meta_t <- rbd_meta %>%
  select(study, estimate, ci.lb, ci.ub, p.value) %>%
  column_to_rownames("study") %>%
  t() %>%
  as.data.frame()

write.csv(rbd_meta, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/rbd_meta.csv")
write.csv(rbd_meta_t, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/rbd_meta_t.csv")
```

# ESS regression in each cohort

```{r}
NBA_EUR$ess_total_score <- as.numeric(NBA_EUR$ess_total_score)
NBA_EUR$AAV_ESS <- as.numeric(NBA_EUR$AAV_ESS)
```

```{r}
# linear regression in 4 cohorts separately

study_keep <- c(
  "PPMI-N", "PROMO-EPIPARK", "PROMO-SFB134", "PPMI-G"
)

lm_ess <- NBA_EUR %>%
  filter(study %in% study_keep) %>%
  group_by(study) %>%
  do(tidy(lm(ess_total_score ~ PRS + AAV_ESS + SEX + PC1 + PC2 + PC3 + PC4 + PC5, data = .))) %>%
  ungroup()

```

## meta analysis

```{r}
ess_meta <- lm_ess %>%
  filter(term == "PRS") %>%
  select(study, estimate, std.error, p.value)

# add 95%CI column
ess_meta <- ess_meta %>%
  mutate(
    ci.lb = estimate - 1.96 * std.error,
    ci.ub = estimate + 1.96 * std.error
  )

# Fixed-effect model
fixed <- rma.uni(yi = estimate,
                     sei = std.error,
                     data = ess_meta,
                     method = "FE")

# Random-effects model
random <- rma.uni(yi = estimate,
                      sei = std.error,
                      data = ess_meta,
                      method = "REML")
					  
writeLines(capture.output(summary(fixed)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/ess_fixed.txt")
writeLines(capture.output(summary(random)),
           "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/ess_random.txt")
```

```{r}
random_row <- tibble(
  study = "Random-effects Model",
  estimate = as.numeric(random$b),
  std.error = random$se,
  p.value = random$pval,
  ci.lb = random$ci.lb,
  ci.ub = random$ci.ub
)
ess_meta <- bind_rows(ess_meta, random_row)

fixed_row <- tibble(
  study = "Fixed-effects Model",
  estimate = as.numeric(fixed$b),
  std.error = fixed$se,
  p.value = fixed$pval,
  ci.lb = fixed$ci.lb,
  ci.ub = fixed$ci.ub
)
ess_meta <- bind_rows(ess_meta, fixed_row)

ess_meta_t <- ess_meta %>%
  select(study, estimate, ci.lb, ci.ub, p.value) %>%
  column_to_rownames("study") %>%
  t() %>%
  as.data.frame()

write.csv(ess_meta, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/ess_meta.csv")
write.csv(ess_meta_t, "/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/results/ess_meta_t.csv")
```

# supplenmentary table

```{r}
demo <- read_excel("/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/dat/demographics_tables2.xlsx")
```

```{r}
demo <- demo %>%
  mutate(age_of_onset = as.numeric(age_of_onset))
```

## sex

```{r}
# distribution
demo %>%
  mutate(
    PHENO_group = case_when(
      PHENO == 1 ~ "control",
      PHENO == 2 ~ "case"
    )
  ) %>%
  group_by(ANCESTRY, PHENO_group) %>%
  summarise(
    total_nonmissing_sex = sum(!is.na(SEX)),
    male_n = sum(SEX == 1, na.rm = TRUE),
    percent_male = 100 * male_n / total_nonmissing_sex,
    .groups = "drop"
  )
```

## age at genotype

```{r}
demo %>%
  group_by(ANCESTRY, PHENO) %>%
  summarise(
    n = n(),
    mean_AGE = mean(AGE, na.rm = TRUE),
    sd_AGE = sd(AGE, na.rm = TRUE)
  ) %>%
  arrange(ANCESTRY, PHENO)
```

## age at onset

```{r}
demo %>%
  filter(!is.na(age_of_onset)) %>%
  group_by(ANCESTRY, PHENO) %>%
  summarise(
    n = n(),
    mean_AAO = mean(age_of_onset, na.rm = TRUE),
    sd_AAO = sd(age_of_onset, na.rm = TRUE)
  ) %>%
  arrange(ANCESTRY, PHENO)
```

# normal distribution

```{r}
df <- read_csv("/home/rstudio/workspace/ws_files/r11/cor_r11/NBA/dat/risk_results_r11.csv")
```

```{r}
# Keep only cases: PHENO = 2
df_case <- df %>%
  mutate(
    PRS = as.numeric(PRS),
    PHENO = as.numeric(PHENO),
    ANCESTRY = as.factor(ANCESTRY)
  ) %>%
  filter(PHENO == 2, !is.na(PRS), !is.na(ANCESTRY))

# Check sample size by ancestry
df_case %>%
  count(ANCESTRY)
```

```{r}
ggplot(df_case, aes(sample = PRS)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(~ ANCESTRY, scales = "free") +
  theme_bw() +
  labs(
    title = "QQ plot of PRS by ancestry among cases only",
    x = "Theoretical quantiles",
    y = "Sample quantiles"
  )
```
